High-level cognition is supported by at least second order dynamic correlations in neural activity patterns

### 2019-08-24 paper/main.tex
working on dimensionality reduction section of methods

\section*{Methods}
Our general approach to comprises four general steps.  First, we
derive a kernel-based approach to computing dynamic pairwise
correlations in a $T$ (timepoints) by $K$ (features) multivariate
timeseries, $\mathbf{X}_0$.  This yields a $T$ by $\mathcal{O}(K^2)$ matrix of
dynamic correlations, $\mathbf{Y}_1$, where each row comprises the upper
triangle of the correlation matrix at a single timepoint, reshaped
into a row vector.  Second, we apply a dimensionality reduction step
to project the matrix of dynamic correlations back onto a
$K$-dimensional space.  This yields a $T$ by $K$ matrix, $\mathbf{X}_1$, that
reflects an approximation of the dynamic correlations reflected in the
original data.  Third, we use repeated applications of the
kernel-based dynamic correlation step to $\mathbf{X}_n$ and the dimensionality
reduction step to the resulting $\mathbf{Y}_{n+1}$ to estimate high-order
dynamnic correlations.  Each application of these steps to a $T$ by
$K$ time series $\mathbf{X}_n$ yields a $T$ by $K$ matrix, $\mathbf{X}_{n+1}$, that
reflects the dynamic correlations between the columns of $\mathbf{X}_n$.  In
this way, we refer to $n$ as the \textit{order} of the timeseries,
where $\mathbf{X}_0$ (order 0) denotes the original data and $\mathbf{X}_n$ denotes
$n^\mathrm{th}$-order dynamic correlations between the columns of
$\mathbf{X}_0$.  Finally, we use a cross-validation--based decoding approach to
evaluate how well information contained in a given order (or weighted
mixture of orders) may be used to decode relevant cognitive states.
If including a given $\mathbf{X}_n$ in the feature set yields higher
classification accuracy on held-out data, we interpret this as
evidence that the given cognitive states are reflected in patterns of
$n^\mathrm{th}$-order correlations.

@@

\subsubsection*{Inter-subject dynamic correlations}
Equation~\ref{eqn:timecorr} provides a means of taking a single
observation matrix, $\mathbf{X}_n$ and estimating the dynamic
correlations from moment to moment, $\mathbf{Y}_{n+1}$.  Suppose that
one has access to a set of multiple observation matrices that reflect
the same phenomenon.  For example, one might collect neuroimaging data
from several experimental participants, as each participant performs
the same task (or sequence of tasks).  Let $\mathbf{X}_n^1$,
$\mathbf{X}_n^2$, ..., $\mathbf{X}_n^P$ reflect the $T$ by $K$
observation matrices ($n = 0$) or reduced correlation matrices
($n > 0$) for each of $P$ participants in an experiment.  We can use
\textit{inter-subject functional connectivity}~\citep[ISFC;
][]{SimoEtal16} to compute the degree of stimulus-driven correlations
reflected in the multi-participant dataset at a given timepoint $t$
using:
\begin{align}
\bar{\mathbf{C}}(t) = M\left(R\left(\frac{1}{2P} \sum_{p=1}^P
  Z\left(Y_n^p(t)\right)^\mathrm{T} + Z\left(Y_n^p(t)\right)\right)\right),
\end{align}
where $M$ extracts and vectorizes the diagonal and upper triangle of a symmetric
matrix, $Z$ is the Fisher $z$-transformation~\citep{Zar10}:
\begin{align}
Z(r) = \frac{\log(1+r) - \log(1-r)}{2}
\end{align}
$R$ is the inverse of $Z$:
\begin{align}
R(z) = \frac{\exp(2z - 1)}{\exp(2z + 1)},
\end{align}
and $\mathbf{Y}_n^p(t)$ denotes the correlation matrix
(Eqn.~\ref{eqn:corr}) between each column of $\mathbf{X}_n^p$ and each
column of the average $\mathbf{X}_n$ from all \textit{other}
participants, $\bar{\mathbf{X}}_n^{ \backslash p}$:
\begin{align}
  \bar{\mathbf{X}}_n^{ \backslash p} = R\left(\frac{1}{P-1}\sum_{q \in
  \backslash p} Z\left( \mathbf{X}_n^q \right) \right),
\end{align}
where $\backslash p$ denotes the set of all participants other than
participant $p$. In this way, the $T$ by $\mathcal{O}(K^2)$
matrix $\bar{\mathbf{C}}$ provides a time-varying extension of the ISFC
approach developed by \cite{SimoEtal16}.

@@

Given a $T$ by $\mathcal{O}(K^2)$ matrix of dynamic correlations,
$\mathbf{Y}_n$, we propose two general approaches to computing a $T$ by $K$
low-dimensional representation of these correlations, $\mathbf{X}_n$.
The first approach uses dimensionality reduction algorithms to project
$\mathbf{Y}_n$ onto a $K$-dimensional space.  The second approach uses
graph-theoretic measures to characterize the relative positions of
each feature ($k \in \left[1, ..., K \right]$) in the network defined
by the correlation matrix at each timepoint. \textbf{JRM STOPPED HERE}

>>>>

\section*{Methods}
Our general approach to comprises four general steps.  First, we
derive a kernel-based approach to computing dynamic pairwise
correlations in a $T$ (timepoints) by $K$ (features) multivariate
timeseries, $\mathbf{X}_0$.  This yields a $T$ by $\mathcal{O}(K^2)$ matrix of
dynamic correlations, $\mathbf{Y}_1$, where each row comprises the upper
triangle of the correlation matrix at a single timepoint, reshaped
into a row vector (this reshaped vector is $(\frac{K^2 - K}{2})$-dimensional).  Second, we apply a dimensionality reduction step
to project the matrix of dynamic correlations back onto a
$K$-dimensional space.  This yields a $T$ by $K$ matrix, $\mathbf{X}_1$, that
reflects an approximation of the dynamic correlations reflected in the
original data.  Third, we use repeated applications of the
kernel-based dynamic correlation step to $\mathbf{X}_n$ and the dimensionality
reduction step to the resulting $\mathbf{Y}_{n+1}$ to estimate high-order
dynamnic correlations.  Each application of these steps to a $T$ by
$K$ time series $\mathbf{X}_n$ yields a $T$ by $K$ matrix, $\mathbf{X}_{n+1}$, that
reflects the dynamic correlations between the columns of $\mathbf{X}_n$.  In
this way, we refer to $n$ as the \textit{order} of the timeseries,
where $\mathbf{X}_0$ (order 0) denotes the original data and $\mathbf{X}_n$ denotes
$n^\mathrm{th}$-order dynamic correlations between the columns of
$\mathbf{X}_0$.  Finally, we use a cross-validation--based decoding approach to
evaluate how well information contained in a given order (or weighted
mixture of orders) may be used to decode relevant cognitive states.
If including a given $\mathbf{X}_n$ in the feature set yields higher
classification accuracy on held-out data, we interpret this as
evidence that the given cognitive states are reflected in patterns of
$n^\mathrm{th}$-order correlations.

@@

\subsubsection*{Dynamic inter-subject functional connectivity (DISFC)}
Equation~\ref{eqn:timecorr} provides a means of taking a single
observation matrix, $\mathbf{X}_n$ and estimating the dynamic
correlations from moment to moment, $\mathbf{Y}_{n+1}$.  Suppose that
one has access to a set of multiple observation matrices that reflect
the same phenomenon.  For example, one might collect neuroimaging data
from several experimental participants, as each participant performs
the same task (or sequence of tasks).  Let $\mathbf{X}_n^1$,
$\mathbf{X}_n^2$, ..., $\mathbf{X}_n^P$ reflect the $T$ by $K$
observation matrices ($n = 0$) or reduced correlation matrices
($n > 0$) for each of $P$ participants in an experiment.  We can use
\textit{inter-subject functional connectivity}~\citep[ISFC;
][]{SimoEtal16} to compute the degree of stimulus-driven correlations
reflected in the multi-participant dataset at a given timepoint $t$
using:
\begin{align}
\bar{\mathbf{C}}(t) = M\left(R\left(\frac{1}{2P} \sum_{p=1}^P
  Z\left(Y_n^p(t)\right)^\mathrm{T} + Z\left(Y_n^p(t)\right)\right)\right),
\end{align}
where $M$ extracts and vectorizes the diagonal and upper triangle of a symmetric
matrix, $Z$ is the Fisher $z$-transformation~\citep{Zar10}:
\begin{align}
Z(r) = \frac{\log(1+r) - \log(1-r)}{2}
\end{align}
$R$ is the inverse of $Z$:
\begin{align}
R(z) = \frac{\exp(2z - 1)}{\exp(2z + 1)},
\end{align}
and $\mathbf{Y}_n^p(t)$ denotes the correlation matrix
(Eqn.~\ref{eqn:corr}) between each column of $\mathbf{X}_n^p$ and each
column of the average $\mathbf{X}_n$ from all \textit{other}
participants, $\bar{\mathbf{X}}_n^{ \backslash p}$:
\begin{align}
  \bar{\mathbf{X}}_n^{ \backslash p} = R\left(\frac{1}{P-1}\sum_{q \in
  \backslash p} Z\left( \mathbf{X}_n^q \right) \right),
\end{align}
where $\backslash p$ denotes the set of all participants other than
participant $p$. In this way, the $T$ by $\frac{K^2 - K}{2}$ DISFC
matrix $\bar{\mathbf{C}}$ provides a time-varying extension of the ISFC
approach developed by \cite{SimoEtal16}.

@@

Given a $T$ by $\frac{K^2 - K}{2}$ matrix of dynamic correlations,
$\mathbf{Y}_n$, we propose two general approaches to computing a $T$
by $K$ low-dimensional representation of these correlations,
$\mathbf{X}_n$.  The first approach uses dimensionality reduction
algorithms to project $\mathbf{Y}_n$ onto a $K$-dimensional space.
The second approach uses graph-theoretic measures to characterize the
relative positions of each feature ($k \in \left[1, ..., K \right]$)
in the network defined by the correlation matrix at each timepoint.

Because $\mathbf{X}_n$ will have the same shape as the original data
$\mathbf{X}_0$, approximating $\mathbf{Y}_n$ with a lower-dimensional
$\mathbf{X}_n$ enables us to estimate high-order dynamic correlations
in a scalable way.  Given a $T$ by $K$ input matrix, the output of
Equation~\ref{eqn:timecorr} requires $\mathcal{O}(TK^2)$ space to
store.  Repeated applications of Equation~\ref{eqn:timecorr} (i.e.,
computing dynamic correlations between the columns of the outputted
dynamic correlation matrix) each require exponentially more space; in
general the $n^\mathrm{th}$-order dynamic correlations of a $T$ by $K$
timeseries occupies $\mathcal{O}(TK^{2n})$ space.  However, when we
approximate the output of Equation~\ref{eqn:timecorr} with a $T$ by
$K$ matrix (as described next), it becomes feasible to compute even
very high-order correlations in high-dimensional data.  Specifically,
approximating the $n^\mathrm{th}$-order dynamic correlations of a $T$
by $K$ timeseries requires only $\mathcal{O}(TK^2)$ space-- the same
as would be required to compute first-order dynamic
correlations. \textbf{JRM STOPPED HERE}
